---
# title: "reference data"
author: "wiesehahn"
date: "2020-07-23"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(# fig.width=12, fig.height=8, 
  echo=FALSE, warning=FALSE, message=FALSE)
```

## Reference Data

```{r plot_ref_init}
library(rgdal)
library(leaflet)
library(raster)
library(formattable)
library(tidyverse)
library(plotly)
library(caret)
library(RColorBrewer)
library(kableExtra)
library(spdplyr)
library(here)

# load data
ref.sol <- readOGR(dsn=here("/data/reference/Solling"), layer = "Training_Solling_2", stringsAsFactors = T, verbose=F)
ref.sol@data$site = as.factor("Solling")

ref.har <- readOGR(dsn=here("/data/reference/Harz"), layer = "Training_Harz_2", stringsAsFactors = T, verbose=F)
ref.har@data$site = as.factor("Harz")

ref.hei <- readOGR(dsn=here("/data/reference/Heide"), layer = "Training_Heide_2", stringsAsFactors = T, verbose=F)
ref.hei@data$site = as.factor("Heide")

ref.all <- rbind(ref.sol,ref.har,ref.hei)


ref.all@data$BA <- plyr::revalue(ref.all@data$BA, c("211"="BU",
                                                    "112"="EI",
                                                    "511"="FI",
                                                    "611"="DGL",
                                                    "711"="KI",
                                                    "811"="LAE"))

                    
                    
ref.all@data <- ref.all@data %>% 
  mutate(BA = fct_relevel(BA, c("BU", "DGL", "FI", "KI", "LAE", "EI")))


# transform
ref <- spTransform(ref.all, CRS("+init=epsg:4326"))

# roi extends
ext.sol <- extent(526000,563000,5716000,5747000)
ext.har <- extent(581000,609780,5715000,5758000)
ext.hei <- extent(580000,594000,5844000,5857000)

# # set same relative extent
# ext.sol <- extent(ref.sol)
# ext.dist <- min(ext.sol@xmax - ext.sol@xmin, ext.sol@ymax - ext.sol@ymin)
# 
# ext.sol@xmax <- ext.sol@xmin + ext.dist
# ext.sol@ymax <- ext.sol@ymin + ext.dist
# 
# ext.har <- extent(ref.har)
# ext.har@xmax <- ext.har@xmin + ext.dist
# ext.har@ymax <- ext.har@ymin + ext.dist
# 
# ext.hei <- extent(ref.hei)
# ext.hei@xmax <- ext.hei@xmin + ext.dist
# ext.hei@ymax <- ext.hei@ymin + ext.dist


# set color palette for leaflet
pal <- colorFactor(c("#bf7013","#a5027c","#4066aa","#94a6b0","#e20030", "#fecc00"), ref@data$BA)

# set color palette for plotly 
pal.ref <- c("#bf7013","#a5027c","#4066aa","#94a6b0","#e20030", "#fecc00") %>%
   setNames(levels(ref@data$BA))

# palette  in the order of reference data levels
pal.class <- c("#bf7013","#a5027c","#4066aa","#94a6b0","#e20030", "#fecc00") 

```

The reference data was sampled in three project areas (Harz, Heide, Solling).
```{r plot_ref_all, fig.cap= "Reference data sampling locations"}
leaflet(ref) %>%
  addProviderTiles(providers$Esri.WorldTopoMap) %>%
  addPolygons( color = ~pal(BA), weight = 1, smoothFactor = 0.5,
              opacity = 1.0, fillOpacity = 0.7,
              highlightOptions = highlightOptions(color = "white", weight = 2,
                                                  bringToFront = TRUE)) %>%
  addLegend("bottomright", pal = pal, values = ~BA,
            title = "Baumart",
            opacity = 1
  )
```


### Number of reference data polygons by tree species
```{r tab_ref_spec, eval=FALSE}
ref@data %>% dplyr::group_by(BA) %>% dplyr::summarise(n=n(), .groups = 'drop') %>% formattable()

```

```{r plot_ref_spec}
ref@data %>% dplyr::group_by(BA) %>% dplyr::summarise(n=n(), .groups = 'drop') %>% 
  plot_ly(labels = ~BA, values = ~n, type = 'pie',
          marker = list(colors = pal.ref,
                      line = list(color = '#FFFFFF', width = 1)))
```


### Number of reference data polygons by site
```{r tab_ref_site, eval=FALSE}

ref@data %>% dplyr::group_by(site) %>% dplyr::summarise(n=n()) %>% formattable()

```

```{r plot_ref_site}

ref@data %>% dplyr::group_by(site) %>% dplyr::summarise(n=n()) %>% 
    plot_ly(labels = ~site, values = ~n, type = 'pie',
          marker = list(colors = brewer.pal(n = 3, name = "Paired"),
                      line = list(color = '#FFFFFF', width = 1)))

```

### Number of reference data polygons by tree species and site
```{r tab_ref_site-spec, eval=FALSE}

ref@data %>% dplyr::group_by(BA, site) %>% dplyr::summarise(n=n()) %>%  spread(site, n) %>% formattable()

```

```{r plot_ref_site-spec}
ref@data %>% dplyr::group_by(BA, site) %>% dplyr::summarise(n=n()) %>%
  plot_ly(x = ~site, y = ~n, type = 'bar', color = ~BA, colors = pal.ref) 
```


-----------------
```{r plot_prob_init}
library(rasterVis)

f <- "G:/B/SG4/wiesehahn/von_philip/RandomForest_Baumarten/rfprob_masked.tif"
if (!file.exists(f)){
  img.class <- raster("G:/B/SG4/wiesehahn/von_philip/RandomForest_Baumarten/S2B_MSIL2A_20170823T103019_N0205_R108_T32UND_20170823T103018_Dsen2_TopCorSlope10K06ndvi_VI_HrzSolHei_int1U_neu.tif")
  img.prob.unmasked <- raster("G:/B/SG4/wiesehahn/von_philip/RandomForest_Baumarten/S2B_MSIL2A_20170823T103019_N0205_R108_T32UND_20170823T103018_Dsen2_TopCorSlope10K06ndvi_VI_HrzSolHei_int1U_neu_rfprob.tif")
  img.prob.masked <- mask(img.prob.unmasked, mask= img.class, filename= "G:/B/SG4/wiesehahn/von_philip/RandomForest_Baumarten/rfprob_masked.tif")
}

img.prob <- raster("G:/B/SG4/wiesehahn/von_philip/RandomForest_Baumarten/rfprob_masked.tif")
# 
# img.prob.sol <- crop(img.prob, ext.sol)
# img.prob.har <- crop(img.prob, ext.har)
# img.prob.hei <- crop(img.prob, ext.hei)

```


### Reference Probability

Classification probabilities were extracted for each pixel inside reference polygons. The extracted values grant insight in classification certainty by tree species and reference site.

```{r plot_ref_prob_init}
#clip <- crop(img.prob, extent(ref.all))

ref.prob.sol <- raster::extract(img.prob, ref.sol, df=TRUE)
ref.prob.sol$ID <- as.factor(ref.prob.sol$ID)
ref.prob.sol <- merge(ref.prob.sol, ref.sol@data ,by="ID", all = TRUE)


ref.prob.har <- raster::extract(img.prob, ref.har, df=TRUE)
ref.prob.har$ID <- as.factor(ref.prob.har$ID)
ref.prob.har <- merge(ref.prob.har, ref.har@data ,by="ID", all = TRUE)


ref.prob.hei <- raster::extract(img.prob, ref.hei, df=TRUE)
ref.prob.hei$ID <- as.factor(ref.prob.hei$ID)
ref.prob.hei <- merge(ref.prob.hei, ref.hei@data ,by="ID", all = TRUE)

ref.prob <- rbind.data.frame(ref.prob.sol, ref.prob.har, ref.prob.hei)

ref.prob$BA <- plyr::revalue(ref.prob$BA, c("211"="BU",
                                                    "112"="EI",
                                                    "511"="FI",
                                                    "611"="DGL",
                                                    "711"="KI",
                                                    "811"="LAE"))

                    
                    
ref.prob <- ref.prob %>% 
  mutate(BA = fct_relevel(BA, c("BU", "DGL", "FI", "KI", "LAE", "EI")))

```

```{r plot_ref_prob, fig.cap= "Classification probability by site and tree species for reference data locations"}
library(plotly)

fig <- plot_ly(ref.prob, x = ~site, y = ~rfprob_masked, color = ~BA, colors= pal.ref, type = "box")

fig <- fig %>% layout(boxmode = "group", yaxis = list(title = "Probability"), xaxis = list(title = FALSE))

fig
```


-----------------


## Mapping Accuracy

### Reference data for validation

Class predictions for all reference pixels were extracted from the model prediction raster. These predictions were thought to be compared with the reference data label to produce an error matrix. The accuracy was expected to be biased since we used part of the reference data for training the model. But instead, all reference data was classified correctly. This might suggest that the model is overfitted to the reference data, performing very well on the reference data but weaker outside.
```{r plot_class_init}

img.class <- raster("G:/B/SG4/wiesehahn/von_philip/RandomForest_Baumarten/S2B_MSIL2A_20170823T103019_N0205_R108_T32UND_20170823T103018_Dsen2_TopCorSlope10K06ndvi_VI_HrzSolHei_int1U_neu.tif")
names(img.class) <- "classification"
```

```{r plot_ref_class_init}
#clip <- crop(img.class, extent(ref.all))

ref.class.sol <- raster::extract(img.class, ref.sol, df=TRUE)
ref.sol@data$rownumber <- as.integer(rownames(ref.sol@data))+1
ref.class.sol <- merge(ref.class.sol, ref.sol@data ,by.x="ID", by.y="rownumber", all = TRUE)


ref.class.har <- raster::extract(img.class, ref.har, df=TRUE)
ref.har@data$rownumber <- as.integer(rownames(ref.har@data))+1
ref.class.har <- merge(ref.class.har, ref.har@data ,by.x="ID", by.y="rownumber", all = TRUE)


ref.class.hei <- raster::extract(img.class, ref.hei, df=TRUE)
ref.hei@data$rownumber <- as.integer(rownames(ref.hei@data))+1
ref.class.hei <- merge(ref.class.hei, ref.hei@data ,by.x="ID", by.y="rownumber", all = TRUE)

ref.class <- rbind.data.frame(ref.class.sol, ref.class.har, ref.class.hei)

# rename columns
colnames(ref.class)[2] <- "classification"

colnames(ref.class)[3] <- "reference"

colnames(ref.class)[1] <- "ID2"

colnames(ref.class)[4] <- "ID"


# revalue classificaton data
ref.class$classification <- as.factor(ref.class$classification)

ref.class$classification <- plyr::revalue(ref.class$classification, c("6"="EI", 
                    "1"="BU",
                    "3"="FI",
                    "2"="DGL",
                    "4"="KI",
                    "5"="LAE"))



ref.class$reference <- plyr::revalue(ref.class$reference, c("211"="BU",
                                                    "112"="EI",
                                                    "511"="FI",
                                                    "611"="DGL",
                                                    "711"="KI",
                                                    "811"="LAE"))
ref.class <- ref.class %>% 
  mutate(classification = fct_relevel(classification, c("BU", "DGL", "FI", "KI", "LAE", "EI")),
         reference = fct_relevel(reference, c("BU", "DGL", "FI", "KI", "LAE", "EI")))

```

**Error Matrix**
```{r plot_ref_class_em, fig.cap= "Error matrix for reference data, which was also used for training the random forest model"}
cm<- confusionMatrix(data = ref.class$classification, reference = ref.class$reference)
cm.df <- as.data.frame.matrix(cm$table)

cm.df %>%
  knitr::kable("html", escape = F, rownames = T,)%>%
  kable_styling("hover", full_width = F) 

```


### National forest inventory data for validation

BWI-plots in Lower Saxony were filtered by certain criteria to serve as validation data. Only plots with a relative tree species proportion of more than 75% in the main canopy layer for one of the classified tree species groups were considered.
Class predictions for all pixels covered by these plots were extracted from the model prediction raster and compared against the inventory data.

```{r plot_bwi_class_init}
# load bwi data
ref_ni <- read.csv(here("/data/reference/bwi/ref_ni.csv"), stringsAsFactors=TRUE)

# ref_ni$BA <- plyr::revalue(ref_ni$BaGr, c("EI"="Eiche", 
#                     "BU"="Buche",
#                     "FI"="Fichte",
#                     "DGL"="Douglasie",
#                     "KI"="Kiefer",
#                     "LAE"="Laerche"))

ref_ni.sp<- SpatialPointsDataFrame(coords = ref_ni[,c("GK3_Rechts", "GK3_Hoch")], data = ref_ni, proj4string = CRS("+init=epsg:31467"))

# transform
ref.bwi <- spTransform(ref_ni.sp, CRS("+init=epsg:4326"))



ref.class.bwi <- raster::extract(img.class, ref.bwi, df=TRUE)

ref.class.bwi <- merge(ref.class.bwi, ref.bwi@data ,by.x="ID", by.y="X", all = TRUE)

# rename columns
colnames(ref.class.bwi)[2] <- "classification"

ref.class.bwi <- ref.class.bwi %>% rename(reference = BaGr)


# revalue classificaton data
ref.class.bwi$classification <- as.factor(ref.class.bwi$classification)

ref.class.bwi$classification <- plyr::revalue(ref.class.bwi$classification, c("6"="EI", 
                    "1"="BU",
                    "3"="FI",
                    "2"="DGL",
                    "4"="KI",
                    "5"="LAE"))

ref.class.bwi <- ref.class.bwi %>% 
  mutate(classification = fct_relevel(classification, c("BU", "DGL", "FI", "KI", "LAE", "EI")),
         reference = fct_relevel(reference, c("BU", "DGL", "FI", "KI", "LAE", "EI")))

ref.class.sp <- sp::merge(ref.bwi, ref.class.bwi, by= c("Tnr", "Enr"))

#mapview(ref.class.sp, zcol="reference", col.regions = pal.ref)

library(spdplyr)
misclassified <- ref.class.sp %>% mutate(result = if_else(classification != reference,"wrong", "right"))
#mapview(misclassified, zcol="result", col.regions = c("green", "red"))
```


**Error Matrix**
```{r plot_bwi_class_em, fig.cap= "Error matrix for inventory data, which was collected 2012 for the NFI"}
cm<- confusionMatrix(data = ref.class.bwi$classification, reference = ref.class.bwi$reference)
cm.df <- as.data.frame.matrix(cm$table)

cm.df %>%
  knitr::kable("html", escape = F, rownames = T,)%>%
  kable_styling("hover", full_width = F) 

```


**Respective Accuracy**
```{r print_bwi_acc}

cm.df <- as.data.frame(t(cm$overall[1:2]))

kable(cm.df) %>% kable_styling(full_width = F)

```


### Federal forest inventory data for validation

BI-plots within the study areas were used as validation data. Plot data was aggregated by species group and basal area was used to determine the dominant tree species per plot. Only plots with a relative basal area proportion of at least 75% for one of the classified tree species groups were considered. Plots with a more heterogeneous species cover and plots dominated by other species groups were not considered.
Class predictions for these plots were extracted from the model prediction raster and compared against the inventory data.

<!-- ```{r plot_bi_class_init} -->
<!-- # load bi data -->
<!-- bi <- read.csv2(here("data/reference/bi/bi2018_nlf_tblDatPh2.csv"), stringsAsFactors=TRUE) -->

<!-- # join species groups -->
<!-- bt_code <- read.csv2(here("data/reference/bestandestyp_ni.csv"), stringsAsFactors=TRUE) -->
<!-- bi <- bi %>% -->
<!--   left_join(bt_code, by = c("DatPh2_BT" = "Ziffer")) -->

<!-- # merge duplicate entries (for different years) -->
<!-- bi <- bi %>% group_by(DatPh2_RW, DatPh2_HW) %>% -->
<!--   slice_max(DatPH2_STJ) %>% -->
<!--   group_by(DatPh2_RW, DatPh2_HW, DatPH2_STJ) %>% -->
<!--   slice_max(DatPh2_BT) %>% -->
<!--   ungroup() -->

<!-- # make spatial -->
<!-- bi.sp<- SpatialPointsDataFrame(coords = bi[,c("DatPh2_RW", "DatPh2_HW")], data = bi, proj4string = CRS("+init=epsg:31467")) -->

<!-- # transform -->
<!-- bi.sp <- spTransform(bi.sp, img.class@crs) -->

<!-- # filter by roi extents -->
<!-- bi_sol.sp <- crop(bi.sp, ext.sol) %>% mutate(site = "solling") -->
<!-- bi_har.sp <- crop(bi.sp, ext.har) %>% mutate(site = "harz") -->
<!-- bi_hei.sp <- crop(bi.sp, ext.hei) %>% mutate(site = "heide") -->

<!-- #merge -->
<!-- bi_roi.sp <- rbind(bi_sol.sp, bi_har.sp, bi_hei.sp) -->

<!-- # extract classification results at bi plots -->
<!-- ref.class.bi <- raster::extract(img.class, bi_roi.sp, sp=TRUE) -->

<!-- # refactor, relevel, compare, ... -->
<!-- ref.class.bi <- ref.class.bi %>% mutate(reference = as.factor(Baumart), -->
<!--                                         site = as.factor(site), -->
<!--                                          classification = fct_recode(as.factor(classification), "EI"="6", "BU"="1","FI"="3","DGL"="2","KI"="4","LAE"="5")) %>% -->
<!--   filter(reference %in% c("BU", "DGL", "FI", "KI", "LAE", "EI")) %>% -->
<!--   mutate(reference = factor(fct_relevel(reference, c("BU", "DGL", "FI", "KI", "LAE", "EI"))), -->
<!--          result = as.factor(if_else(classification != reference,"incorrect", "correct")), -->
<!--          label = paste(sep ="  ",'Reference: ', as.character(reference), 'Classification: ',as.character(classification)) -->
<!-- ) -->


<!-- ref.class.bi.df <- ref.class.bi@data -->

<!-- #calculate species composition -->
<!-- species.composition <- ref.class.bi.df %>% group_by(site, reference) %>% summarise(Anteil = n()) %>% mutate(Anteil = round(Anteil/sum(Anteil)*100, 1)) %>% spread(site, Anteil)  -->

<!-- write.csv(species.composition, file= here("data/reference/bi/bi2018_nlf_speciescomposition_by_roi.csv")) -->

<!-- #calculate results by stand type -->
<!-- standtype.result <- ref.class.bi.df %>% filter(result %in% c("correct", "incorrect")) %>% group_by(Bestandestyp, result) %>% summarise(Anteil = n()) %>% mutate(Anteil = round(Anteil/sum(Anteil)*100, 1)) %>% spread(result, Anteil)  -->

<!-- write.csv(standtype.result, file= here("data/reference/bi/bi2018_nlf_classificationresult_by_bestandestyp.csv")) -->

<!-- # filter for pure stand types -->
<!-- ref.class.bi.pure <- ref.class.bi.df %>% filter(Bestandestyp %in% c("Bu rein","Dgl rein", "Fi rein", "Ki rein", "Lae rein", "Ei rein")) -->
<!-- ``` -->

<!-- **Map overview** -->
<!-- ```{r plot_ref_bi, fig.cap= "Reference data sampling locations"} -->
<!-- pal.result <- colorFactor(c("red", "grey"), levels = c("incorrect", "correct")) -->

<!-- # transform -->
<!-- ref.bi <- spTransform(ref.class.bi, CRS("+init=epsg:4326")) -->

<!-- ref.bi %>%  -->
<!-- leaflet() %>% -->
<!--   addProviderTiles(providers$Esri.WorldImagery) %>% -->
<!--   addCircleMarkers( -->
<!--     radius = 5, -->
<!--     color = ~pal.result(result), -->
<!--     stroke = FALSE, fillOpacity = 0.5, -->
<!--     label = ~label -->
<!--   ) %>% -->
<!--   addLegend("bottomright", pal = pal.result, values = ~result, -->
<!--             title = "Classification result", -->
<!--             opacity = 1 -->
<!--   ) -->
<!-- ``` -->

<!-- **Error Matrix** -->
<!-- ```{r plot_bi_class_em, fig.cap= "Error matrix for inventory data, which was collected 2012 for the NFI"} -->
<!-- cm<- confusionMatrix(data = ref.class.bi.df$classification, reference = ref.class.bi.df$reference) -->
<!-- cm.df <- as.data.frame.matrix(cm$table) -->

<!-- cm.df %>% -->
<!--   knitr::kable("html", escape = F, rownames = T,)%>% -->
<!--   kable_styling("hover", full_width = F)  -->

<!-- ``` -->


<!-- **Respective Accuracy** -->
<!-- ```{r print_bi_acc} -->

<!-- cm.df <- as.data.frame(t(cm$overall[1:2])) -->

<!-- kable(cm.df) %>% kable_styling(full_width = F) -->

<!-- ``` -->


```{r plot_bi_class_init}

## calculate dominant species per plot basd on basal area

# load bi data
bi.baum <- read.csv2(here("data/reference/bi/bi2017_nlf_tblDatPh2_Vorr.csv"), stringsAsFactors=TRUE)

colnames(bi.baum) <- gsub('DatPh2_','',colnames(bi.baum))
colnames(bi.baum) <- gsub('Vorr_','',colnames(bi.baum))

# join species groups
spec_code <- read.csv2(here("data/reference/ba_code_ni_bdat.csv"), stringsAsFactors=TRUE)
bi.baum <- bi.baum %>%
  left_join(spec_code %>% select(id_ni, gruppe), by = c("BA" = "id_ni"))

# calculate species shares per plot (by basal area, crown projection)
bi_plots <- bi.baum %>%
  filter(BestSchicht == 1) %>%
  mutate(ba = pi*(0.5*BHD)^2) %>%
  group_by(KSPNr) %>%
  mutate(ba_plot = sum(ba)) %>%
  ungroup() %>%
  group_by(KSPNr, gruppe) %>%
  summarise(ba = sum(ba)/unique(ba_plot)) %>%
  ungroup()

# species per plot with max basal area
bi_spec <- bi_plots %>%
  group_by(KSPNr) %>%
  slice_max(ba) %>%
  ungroup()

# filter to relevant species groups (leave out SLN and SLB) reorder factors and filter to stands with dominant species
bi_spec <- bi_spec %>%
  filter(gruppe %in% c("BU", "DGL", "FI", "KI", "LAE", "EI")) %>%
  droplevels() %>%
  mutate(species_ba = fct_relevel(gruppe, c("BU", "DGL", "FI", "KI", "LAE", "EI"))) %>%
  filter(ba >= 0.75) %>%
  select(KSPNr, species_ba)


## calculate dominant species based on forest type

bi.plot <- read.csv2(here("data/reference/bi/bi2017_nlf_tblDatPh2.csv"), stringsAsFactors=TRUE)

colnames(bi.plot) <- gsub('DatPh2_','',colnames(bi.plot))

# join species groups
bt_code <- read.csv2(here("data/reference/bestandestyp_ni.csv"), stringsAsFactors=TRUE)
bi.plot <- bi.plot %>% 
  left_join(bt_code, by = c("BT" = "Ziffer")) %>%
  filter(Baumart %in% c("BU", "DGL", "FI", "KI", "LAE", "EI"))%>%
  droplevels()%>%
  mutate(species_ft = fct_relevel(Baumart, c("BU", "DGL", "FI", "KI", "LAE", "EI"))) %>%
  select(KSPNr, RW, HW, Bestandestyp, species_ft)


## merge data
  
bi <- bi_spec %>%
  inner_join(bi.plot, by = c("KSPNr")) 

# make spatial
bi.sp<- SpatialPointsDataFrame(coords = bi[,c("RW", "HW")], data = bi, proj4string = CRS("+init=epsg:31467"))

# transform
bi.sp <- spTransform(bi.sp, img.class@crs)

# filter by roi extents
bi_sol.sp <- crop(bi.sp, ext.sol) %>% mutate(site = "solling")
bi_har.sp <- crop(bi.sp, ext.har) %>% mutate(site = "harz")
bi_hei.sp <- crop(bi.sp, ext.hei) %>% mutate(site = "heide")

#merge
bi_roi.sp <- rbind(bi_sol.sp, bi_har.sp, bi_hei.sp)

# extract classification results at bi plots
ref.class.bi <- raster::extract(img.class, bi_roi.sp, sp=TRUE)

# refactor, relevel, compare, ...
ref.class.bi <- ref.class.bi %>% mutate(reference = as.factor(species_ba),
                                        site = as.factor(site),
                                         classification = fct_recode(as.factor(classification), "EI"="6", "BU"="1","FI"="3","DGL"="2","KI"="4","LAE"="5")) %>%
  mutate(result = as.factor(if_else(classification != reference,"incorrect", "correct")),
         label = paste(sep ="  ",'Reference: ', as.character(reference), 'Classification: ',as.character(classification))
)


ref.class.bi.df <- ref.class.bi@data

#calculate species composition
species.composition <- ref.class.bi.df %>% group_by(site, reference) %>% summarise(Anteil = n()) %>% mutate(Anteil = round(Anteil/sum(Anteil)*100, 1)) %>% spread(site, Anteil) 
 
write.csv(species.composition, file= here("data/reference/bi/bi2017_nlf_speciescomposition_by_roi_ba-greater-075.csv"))

#calculate results by stand type
standtype.result <- ref.class.bi.df %>% filter(result %in% c("correct", "incorrect")) %>% group_by(Bestandestyp, result) %>% summarise(Anteil = n()) %>% mutate(Anteil = round(Anteil/sum(Anteil)*100, 1)) %>% spread(result, Anteil) 

write.csv(standtype.result, file= here("data/reference/bi/bi2017_nlf_classificationresult_by_bestandestyp_ba-greater-075.csv"))

# filter for pure stand types
ref.class.bi.pure <- ref.class.bi.df %>% filter(Bestandestyp %in% c("Bu rein","Dgl rein", "Fi rein", "Ki rein", "Lae rein", "Ei rein"))

```

**Map overview**
```{r plot_ref_bi, fig.cap= "Reference data sampling locations"}
pal.result <- colorFactor(c("red", "grey"), levels = c("incorrect", "correct"))

# transform
ref.bi <- spTransform(ref.class.bi, CRS("+init=epsg:4326"))

ref.bi %>% 
leaflet() %>%
  addProviderTiles(providers$Esri.WorldImagery) %>%
  addCircleMarkers(
    radius = 5,
    color = ~pal.result(result),
    stroke = FALSE, fillOpacity = 0.5,
    label = ~label
  ) %>%
  addLegend("bottomright", pal = pal.result, values = ~result,
            title = "Classification result",
            opacity = 1
  )
```

**Error Matrix**
```{r plot_bi_class_em, fig.cap= "Error matrix for inventory data, which was collected 2012 for the NFI"}
cm<- confusionMatrix(data = ref.class.bi.df$classification, reference = ref.class.bi.df$reference)
cm.df <- as.data.frame.matrix(cm$table)

cm.df %>%
  knitr::kable("html", escape = F, rownames = T,)%>%
  kable_styling("hover", full_width = F) 

```


**Respective Accuracy**
```{r print_bi_acc}

cm.df <- as.data.frame(t(cm$overall[1:2]))

kable(cm.df) %>% kable_styling(full_width = F)

```
